---
title: "shankarz.TexST package"
author: "Natarajan Shankar"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{"summary"}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Package Description  
* Package: shankarz.TexST
* Type: Package
* Title: A Twitter Recommendation Engine
* Version: 1.0
* Date: 2015-03-18
* Author: Natarajan Shankar
* Maintainer: Natarajan Shankar <shankarz@stanford.edu>
* Description: shankarz.TexST : A tweet recommendation engine
* Depends: R (>= 2.10)
* License: GPL (>= 2)
* LazyData: true
* Imports: RCurl, ROAuth, plyr, slam, tm, twitteR, stringr
* Suggests: testthat, knitr
* VignetteBuilder: knitr



## shankarz.TexST - Abstract
The **shankarz.TexST** package implements an information retrieval and processing system based upon Twitter tweets. In social engineering, data exchanged between people carries information that can then be broken down and sorted. Twitter is one such source of inter-people information. Sorted data from Twitter can then be used to help make smart, correlative future decisions. The use of past Twitter data to make future projections represents the crux of this project. 

Given a collective set of tweets that correspond to a specific search string in a specific geography in a specific time frame:

**shankarz.TexST** uses the TF-IDF algorithm to prioritize the ASCII tags contained in the collective set and forms a prioritized dataset.
  2). The prioritization of the ASCII tags relies upon the well known TF-IDF algorithm (**Reference: http://en.wikipedia.org/wiki/Tf–idf**)
  3). Given a new search string and the pre-existing prioritized ASCII tags, shankarz.TexST searches the prioritized dataset for relevant tweets.
  
The utility of this package is that a user with a general idea of what they are looking for in Twitter can use a very general search string, feed it as a filter into a large prioritized dataset, and then be able to retrieve a list of the top N relevant tweets. This project, conducted for as part of the requirements for the Stats290 curriculum, is a rudimentary foray into the area of filtered search. *While the prioritized datasets may not exist today, it is expected that data summarization and filtered search will become an important aspect in future years*.


## Visible functions available in shankarz.TexST package

1. ** twitterAuthenticate()                                  Set up and complete Twitter authentication  **

    ** Description**
                twitterAuthenticate()
                Authenticates user session with Twitter

    **Usage**
                twitterAuthenticate()
    **Details**
                Prerequisite task (to be completed at dev.twitter.com) 
                The dev.twitter.com application control panel provides the ability to generate an OAuth access 
                token for the owner of this application. Visit the -My applications- page by navigating to 
                appsdottwitterdotcom, or hover over the profile image in the top right hand corner of the twitter 
                site and select -My applications-. 

                The Applications page contains a list of the applications, along with a button to create a new 
                application. Select the application you wish to generate a token for, and click on its title. Then 
                Click on the -Create my access token- button, and an authorized access token and secret will be 
                generated for this account and the current application. 

                Once the access tokens have been generated, they are to be into this function. Knowing the tokens, 
                the shankarz.TexST application will subsequently make requests on behalf of this R session.

                Once the prequisite task is complete, twitterAuthenticate will succeed in accessing the Twitter server 
                with the pre-generated access token and secret: 

    **Author**
              Natarajan Shankar
              
    ** References**
              http://dev.twitter.com
              
    ** Example function call**
              ```
              twitterAuthenticate()
                ```

1. **  twitterBuildDatabase()                Get tweets from Twitter, given user credentials and tokens**

    **Description**

      twitterBuildDatabase() 
      
      Gets tweets from within an area within miles of a location specified by lattitude and longitude

    **Usage**

      twitterBuildDatabase(search.string = "", lat = "", long = "", miles = "", since = "", numberOfTweets = 1)
      
    **Arguments**

      1. search.string  An optional search string to query Twitter.
      2. lat            A string of latitude numerics.
      3. long            A string of longitude numeric.
      4. miles	        The number of miles for radius of search (as string).
      5. since          The start date for the Twitter search ("yyyy-mm-dd").
      6. numberOfTweets	Number of tweets requested from Twitter.

  **Details**

      This function conducts a search that pulls tweets that originated within specified miles of a location (lat, long) that have been 
      tweeted after the date specified by the since parameter. Optionally, this function also accepts a search string for the query.

  **Value**

      A data.frame of twitter tweets.

  **Example function call**
```
    message("To run twitterBuildDatabase, it is mandatory to have have run TwitterAuthenticate()")
    twitterAuthenticate()

    twitterBuildDatabase(search.string = "Giants", lat="37.794108", long="-122.39511",
           miles="15", since="2014-10-01", numberOfTweets = 100)
 ```

1. # generateTweetRecommendations()                                Given a new tweet, return n recommended tweets.**


    **Description**

      generateTweetRecommendations()
      * Recommends tweets in the corpus given a seed tweet.

    **Usage**

      generateTweetRecommendations(tweetDF, tweet = NULL, num_recs = 10)

    **Arguments**

      1. tweetDF  
      * A data.frame returned from the prior step that called twitterBuildDatabase().

      2. tweet	
      * The seed string for which recommendations will be returned for.

      3. num_recs	
      * The number of recommendations requested.

    **Details**

      This function constructs a DocumentTermMatrics (reference: package tm) that is weighted by TF-IDF in order to 
      show similar words to any given tweet. Given a seed tweet, this function adds the tweet to the corpus, calculates 
      TF-IDF on the entire set, and recommends n similar tweets based on text similarity.

    **Value**

      tweets recommended for the seed tweet.

    **Example function call**

      generateTweetRecommendations(sfTweets, tweet="sourdough", num_recs=10)

    **Author**
      Natarajan Shankar
              
    ** References**
      http://dev.twitter.com


## Running the entire code sequence in 3 easy function calls

```{r}
# Load the needed library
library(shankarz.TexST)
```
```
twitterAuthenticate()
```
```
df <- twitterBuildDatabase("sf", lat="37.7941",
                       long="-122.395",
                       miles="15",
                       since="2015-01-01",
                       numberOfTweets=500)
```


```{r}
seedTweet <- "food at the wharf"
generateTweetRecommendations(sfTweets, seedTweet)

seedTweet <- "downtown homeless"
generateTweetRecommendations(sfTweets, seedTweet)
```
